### 2020-05-19 submission/supplementary.Rmd
Optimized front page.

---
title: "**Supplementary Information**"
subtitle: "**Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header_supplementary.tex
fontsize: 12pt
geometry: margin=1.0in
csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
bibliography: references.bib
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE, message=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)

library("kableExtra")
```

\vspace{20mm}
Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^

1\. Center for Marine Research, Ruđer Bošković Institute, Croatia

2\. Department of Functional and Evolutionary Ecology, University of Vienna, Austria

3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands

^$*$^To whom correspondence should be addressed:

Marino Korlević

G. Paliaga 5, 52210 Rovinj, Croatia

Tel.: +385 52 804 768

Fax: +385 52 804 780

e-mail: marino.korlevic@irb.hr

Running title: Seasonal dynamics of epiphytic communities

\sisetup{mode=text}
\setlength\parindent{24pt}

## Supplementary Figures

```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C. nodosa} [Mixed Settlement] and \\textit{C. cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
include_graphics("../results/figures/rarefaction.jpg")
```

```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon Diversity Index and Inverse Simpson Index of bacterial and archaeal communities from the surfaces of macrophytes (\\textit{C. nodosa} [Mixed Settlement] and \\textit{C. cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
include_graphics("../results/figures/calculators.jpg")
```

```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of macrophytes (\\textit{C. nodosa} [Mixed Settlement] and \\textit{C. cylindracea} [Mixed and Monospecific Settlement]) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
include_graphics("../results/figures/chloroplast_bar_plot.jpg")
```

## Supplementary Table

```{r nseq_notus}
# Loading OTU/sample data
shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")

# Calculating the number of sequences and OTUs in each sample
nseq_notus <- shared %>%
  mutate(nseq=rowSums(select(., starts_with("Otu")))) %>%
  mutate(notus=rowSums(select(., starts_with("Otu"))!=0)) %>%
  select(Group, nseq, notus)

# Loading metadata 
metadata <- read_tsv("data/raw/metadata.csv")

# Joining metadata with the number of sequences and OTUs
invisible(Sys.setlocale(locale="en_GB.utf8"))
metadata_nseq_notus <- inner_join(metadata, nseq_notus, by=c("ID"="Group")) %>%
  select(ID, station, date, season, nseq, notus) %>%
  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
  mutate(date=format(date, "%d %B %Y")) %>%
  mutate(date=str_replace(date, "^0", "")) %>%
  arrange(match(station, c("F", "FCyM", "FCaM", "FCa"))) %>%
  mutate(station=str_replace(station, "^F$", "Seawater")) %>%
  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Mixed)")) %>%
  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
  mutate(ID=str_replace(ID, "^23_1$", "23a")) %>%
  mutate(ID=str_replace(ID, "^23_2$", "23b")) %>%
  mutate(nseq=format(nseq, big.mark=","), notus=format(notus, big.mark=","))

# Generating the table
kable(metadata_nseq_notus, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Sample ID", "Community Type", "Date", "Season", "No. of Sequences", "No. of OTUs"), caption="Sample ID, Community Type, Sampling Date and Season, No. of Sequences and No. of OTUs of each sample. No. of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\\label{nseq_notus}", align="c") %>%
  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width="6em")
```