### 2021-02-23 submission/manuscript.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine
Macrophyte Surfaces}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}
\renewcommand{\linenumberfont}{\normalfont\tiny}

\usepackage{pdfpages}

\usepackage{indentfirst}

\usepackage[labelsep=period, labelfont=bf]{caption}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\figurename}{Fig.}
\captionsetup{justification=raggedright,singlelinecheck=false}

\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

\usepackage{siunitx}
\DeclareSIUnit\molar{\mole\per\cubic\deci\metre}
\DeclareSIUnit\Molar{\textsc{m}}
\DeclareSIUnit\cells{\text{cells}}

\usepackage{caption}
\captionsetup{justification=justified}

\usepackage{float}

\usepackage{xr}
\externaldocument[supp-]{supplementary}

\usepackage{txfonts}

\renewcommand{\figureautorefname}{Fig.}

\usepackage{microtype}

\usepackage{setspace}

\begin{document}
\maketitle

\vspace{20mm}

Marino Korlević\textsuperscript{1\(*\)}, Marsej
Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}

1. Center for Marine Research, Ruđer Bošković Institute, Croatia

2. Department of Functional and Evolutionary Ecology, University of
Vienna, Austria

3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
Netherlands Institute for Sea Research, Utrecht University, The
Netherlands

\textsuperscript{\(*\)}To whom correspondence should be addressed:

Marino Korlević

G. Paliaga 5, 52210 Rovinj, Croatia

Tel.: +385 52 804 768

Fax: +385 52 804 780

e-mail:
\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}

Running title: Seasonal dynamics of epiphytic communities

\newpage
\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}
\doublespacing

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Surfaces of marine macrophytes are inhabited by diverse microbial
communities. Most studies focusing on epiphytic communities of
macrophytes did not take into account temporal changes or applied low
sampling frequency approaches. The seasonal dynamics of epiphytic
microbial communities was determined in a meadow of \emph{Cymodocea
nodosa} invaded by \emph{Caulerpa cylindracea} and in a monospecific
settlement of \emph{Caulerpa cylindracea} at monthly intervals. For
comparison the ambient prokaryotic picoplankton community was also
characterized. At the OTU level, the microbial community composition
differed between the ambient water and the epiphytic communities
exhibiting host-specificity. Also, successional changes were observed
connected to the macrophyte growth cycle. Taxonomic analysis, however,
showed similar high rank groups in the ambient water and the epiphytic
communities, with the exception of \emph{Desulfobacterota}, which were
only found on \emph{Caulerpa cylindracea}. \emph{Cyanobacteria} showed
seasonal changes while other high rank taxa were present throughout the
year. Phylogenetic groups present throughout the year constituted most
of the sequences, while less abundant taxa showed seasonal patterns
connected to the macrophyte growth cycle. Taken together, epiphytic
microbial communities of the seagrass \emph{Cymodocea nodosa} and the
macroalga \emph{Caulerpa cylindracea} appear to be host-specific and
contain taxa that undergo successional changes.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Marine macrophytes (seagrasses and macroalgae) are important ecosystem
engineers forming close associations with microorganisms belonging to
all three domains of life (Egan et al., 2013; Tarquinio et al., 2019).
Microbes can live within macrophyte tissue as endophytes or can form
epiphytic communities on surfaces of leaves and thalli (Egan et al.,
2013; Hollants et al., 2013; Tarquinio et al., 2019). Epiphytic and
endophytic microbial communities exhibit a close functional relationship
with the macrophyte host. It was proposed that this close relationship
constitutes a holobiont, an integrated community where the macrophyte
organism and its symbiotic partners support each other (Margulis, 1991;
Egan et al., 2013; Tarquinio et al., 2019).

Biofilms of microbial epiphytes can contain diverse taxonomic groups and
harbor cell abundances from 10\textsuperscript{2} to
10\textsuperscript{7} \si{\cells\per\cm\squared} (Armstrong et al.,
2000; Bengtsson et al., 2010; Burke et al., 2011b). In such an
environment a number of positive and negative interactions between the
macrophyte and the colonizing microorganisms have been described (Egan
et al., 2013; Hollants et al., 2013; Tarquinio et al., 2019).
Macrophytes can promote growth of associated microbes by nutrient
exudation, while in return microorganisms may support macrophyte
performance through improved nutrient availability, phytohormone
production and protection from toxic compounds, oxidative stress,
biofouling organisms and pathogens (Egan et al., 2013; Hollants et al.,
2013; Tarquinio et al., 2019). Besides these positive interactions,
macrophytes can negatively impact the associated microbes such as
pathogenic bacteria by producing reactive oxygen species and secondary
metabolites (Egan et al., 2013; Hollants et al., 2013; Tarquinio et al.,
2019).

All these ecological roles are carried out by a taxonomically diverse
community of microorganisms. At higher taxonomic ranks a core set of
macrophyte epiphytic taxa was described consisting mainly of
\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria},
\emph{Desulfobacterota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
\emph{Actinobacteriota}, \emph{Firmicutes}, \emph{Planctomycetota},
\emph{Chloroflexi} and \emph{Verrucomicrobiota} (Crump and Koch, 2008;
Tujula et al., 2010; Lachnit et al., 2011). In contrast, at lower
taxonomic ranks host specific microbial communities were found (Lachnit
et al., 2011; Roth-Schulze et al., 2016). Recently, it was shown that
even different morphological niches within the same alga had a higher
influence on the composition of the bacterial community than
biogeography or environmental factors (Morrissey et al., 2019). While
the microbial community composition varies between host species,
metagenomic analyses revealed that the majority of the microbial
functions are conserved (Burke et al., 2011a; Roth-Schulze et al.,
2016). This discrepancy between the microbial taxonomic and functional
composition might be explained by the lottery hypothesis. It postulates
that an initial random colonization step takes places from a set of
functionally equivalent taxonomic groups resulting in taxonomically
different epiphytic communities sharing a core set of functional genes
(Burke et al., 2011a; Roth-Schulze et al., 2016). In addition, some of
the variation in the reported data could be attributed to different
techniques used in these studies, such as different protocols for
epiphytic cell detachment and/or DNA isolation, as no standard protocol
has been yet established to study epiphytic communities (Ugarelli et
al., 2019; Korlević et al., submitted).

The majority of studies describing macrophyte epiphytic microbial
communities did not include possible seasonal changes (Crump and Koch,
2008; Lachnit et al., 2009; Burke et al., 2011b; Roth-Schulze et al.,
2016; Ugarelli et al., 2019). If seasonal changes were taken into
account, low temporal frequency, applied methodologies and/or limited
number of analysed host species did not allow a high taxonomic
resolution (Tujula et al., 2010; Lachnit et al., 2011; Michelou et al.,
2013; Miranda et al., 2013; Mancuso et al., 2016). In the present study
we describe the seasonal dynamics of bacterial and archaeal communities
on the surfaces of the seagrass \emph{Cymodocea nodosa} and siphonous
macroalgae \emph{Caulerpa cylindracea} determined on a mostly monthly
scale. Bacterial and archaeal epiphytes were sampled in a meadow of
\emph{C. nodosa} invaded by the invasive \emph{C. cylindracea} and in a
locality of only \emph{C. cylindracea} located in the proximity of the
seagrass meadow. For comparison, the microbial community of the ambient
seawater was also characterized.

\newpage

\hypertarget{materials-and-methods}{%
\subsection{Materials and methods}\label{materials-and-methods}}

\hypertarget{sampling}{%
\subsubsection{Sampling}\label{sampling}}

Sampling was performed in the Bay of Funtana, northern Adriatic Sea
(\ang{45;10;39} N, \ang{13;35;42} E). Leaves and thalli were sampled in
a meadow of \emph{Cymodocea nodosa} invaded by the invasive
\emph{Caulerpa cylindracea} (mixed settlement) and in a monospecific
settlement of \emph{Caulerpa cylindracea} located in the proximity of
the meadow at approximately monthly intervals from December 2017 to
October 2018 (\autoref{supp-nseq_notus}). Leaves and thalli were
collected by diving and transported to the laboratory in containers
placed on ice and filled with seawater collected at the sampling site.
Upon arrival to the laboratory, \emph{C. nodosa} leaves were cut into
sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea} thalli were cut
into 5 -- 8 \si{\cm} long sections. Leaves and thalli were washed three
times with sterile artificial seawater (ASW) to remove loosely attached
microbial cells. Ambient seawater was collected in 10 \si{\l} containers
by diving and transported to the laboratory where 10 -- 20 \si{\l} were
filtered through a 20 \si{\um} net. The filtrate was further
sequentially filtered through 3 \si{\um} and 0.2 \si{\um} polycarbonate
membrane filters (Whatman, United Kingdom) using a peristaltic pump.
Filters were briefly dried at room temperature and stored at \num{-80}
\si{\degreeCelsius}. Seawater samples were also collected approximately
monthly from July 2017 to October 2018.

\hypertarget{dna-isolation}{%
\subsubsection{DNA isolation}\label{dna-isolation}}

DNA from surfaces of \emph{C. nodosa} and \emph{C. cylindracea} was
isolated using a previously modified and adapted protocol that allows
for a selective epiphytic DNA isolation (Massana et al., 1997; Korlević
et al., submitted). Briefly, leaves and thalli are incubated in a lysis
buffer and treated with lysozyme and proteinase K. Following the
incubations, the mixture containing lysed epiphytic cells was separated
from the leaves and thalli and extracted using phenol-chloroform.
Finally, the extracted DNA was precipitated using isopropanol. DNA from
seawater picoplankton was extracted from 0.2 \si{\um} polycarbonate
filters according to Massana et al. (1997) with a slight modification.
Following the phenol-chloroform extraction, 1/10 of chilled 3
\si{\Molar} sodium acetate (pH 5.2) was added. DNA was precipitated by
adding 1 volume of chilled isopropanol, incubating the mixtures
overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 20,000 ×
\emph{g} and 4 \si{\degreeCelsius} for 21 \si{\minute}. The pellet was
washed twice with 500 \si{\ul} of chilled 70 \si{\percent} ethanol and
centrifuged after each washing step at 20,000 × \emph{g} and 4
\si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were re-suspended
in 50 -- 100 \si{\ul} of deionized water.

\hypertarget{illumina-16s-rrna-sequencing}{%
\subsubsection{Illumina 16S rRNA
sequencing}\label{illumina-16s-rrna-sequencing}}

Illumina MiSeq sequencing of the V4 region of the 16S rRNA gene was
performed as described previously (Korlević et al., submitted). The V4
region of the 16S rRNA gene was amplified using a two-step PCR
procedure. In the first PCR, the 515F
(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
Project (\url{https://earthmicrobiome.org/protocols-and-standards/16s/})
were used (Caporaso et al., 2012; Apprill et al., 2015; Parada et al.,
2016). These primers contained on their \(5'\) end a tagged sequence.
Purified PCR products were sent for Illumina MiSeq sequencing at IMGM
Laboratories, Martinsried, Germany. Prior to sequencing at IMGM, the
second PCR amplification of the two-step PCR procedure was performed
using primers targeting the tagged region incorporated in the first PCR.
In addition, these primers contained adapter and sample-specific index
sequences. Beside samples, a positive and negative control for each
sequencing batch was sequenced. The negative control comprised PCR
reactions without DNA template, while for a positive control a mock
community composed of evenly mixed DNA material originating from 20
bacterial strains (ATCC MSA-1002, ATCC, USA) was used. Sequences
obtained in this study have been deposited in the European Nucleotide
Archive (ENA) at EMBL-EBI under accession number PRJEB37267.

\hypertarget{sequence-analysis}{%
\subsubsection{Sequence analysis}\label{sequence-analysis}}

Obtained sequences were analysed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss et al., 2009) according to the MiSeq Standard
Operating Procedure (MiSeq SOP; \url{https://mothur.org/wiki/MiSeq_SOP})
(Kozich et al., 2013) and recommendations provided by the Riffomonas
project to enhance data reproducibility
(\url{http://www.riffomonas.org/}). For alignment and classification of
sequences the SILVA SSU Ref NR 99 database (release 138;
\url{http://www.arb-silva.de}) was used (Quast et al., 2013; Yilmaz et
al., 2014). Pipeline data processing and visualization was done using R
(version 3.6.0) (R Core Team, 2019) combined with packages vegan
(version 2.5-6) (Oksanen et al., 2019), tidyverse (version 1.3.0)
(Wickham et al., 2019) and multiple other packages ({\textbf{???}};
Neuwirth, 2014; Xie, 2014, 2015, 2019; Wilke, 2018; Xie et al., 2018;
Allaire et al., 2019; Zhu, 2019). The detailed analysis procedure
including the R Markdown file are available in the GitHub repository
(\url{https://github.com/MicrobesRovinj/Korlevic_EpiphyticDynamics_FrontMicrobiol_2021}).
Based on the ATCC MSA-1002 mock community included in the analysis an
average sequencing error rate of 0.01 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich et al., 2013; Schloss et al., 2016). In
addition, the negative controls processed together with the samples
yielded on average only 2 sequences after sequence quality curation.

\newpage

\hypertarget{results}{%
\subsection{Results}\label{results}}

Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
sequences after quality curation and exclusion of sequences without
known relatives (no relative sequences) and eukaryotic, chloroplast and
mitochondrial sequences (\autoref{supp-nseq_notus}). A total of 35
samples originating from epiphytic archaeal and bacterial communities
associated with surfaces of the seagrass \emph{C. nodosa} and the
macroalga \emph{C. cylindracea} were analysed. In addition, 18 samples
(one of the samples was sequenced twice) originating from the ambient
seawater were also processed for comparison. The number of reads per
sample ranged between 8,408 and 77,463 sequences
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off as commonly observed in
high-throughput 16S rRNA amplicon sequencing approaches
(\autoref{supp-rarefaction}). Following quality curation and exclusion
of sequences as mentioned above reads were clustered into 28,750
different OTUs at a similarity level of 97 \si{\percent}. Read numbers
were normalized to the minimum number of sequences (8,408,
\autoref{supp-nseq_notus}) through rarefaction resulting in 17,025
different OTUs with 347 to 1,977 OTUs per sample
(\autoref{supp-calculators}). To determine seasonal changes in richness
and diversity the observed number of OTUs, Chao1, ACE, Exponential
Shannon and Inverse Simpson (Jost, 2006) were calculated after
normalization through rarefaction. Generally, richness estimators and
diversity indices showed similar trends. On average, higher values were
found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,683.5 ±
128.5 OTUs{]} and monospecific {[}Number of OTUs, 1,729.9 ± 145.8
OTUs{]}) than for \emph{C. nodosa} (Number of OTUs, 1,052.2 ± 211.4
OTUs) and lowest values were obtained for the microbial community of the
ambient seawater (Number of OTUs, 529.2 ± 146.8 OTUs)
(\autoref{supp-calculators}). Seasonal changes did not reveal such large
dissimilarities. \emph{C. nodosa} communities showed a slow increase
towards the end of the study, while \emph{C. cylindracea} (mixed and
monospecific) communities were characterized by slightly higher values
in spring and summer than in autumn and winter
(\autoref{supp-calculators}).

To determine the proportion of shared archaeal and bacterial OTUs and
communities sampled in different environments the Jaccard's Similarity
Coefficient on presence-absence data and Bray-Curtis Similarity
Coefficient, respectively, were calculated. Coefficients were determined
after normalization through rarefaction and binning of samples from the
particular environment. The highest proportion of shared OTUs and
community was found between mixed and monospecific \emph{C. cylindracea}
(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
calculated between ambient seawater and epiphytic communities
(\autoref{matrix}). Shared proportion between \emph{C. nodosa} and
\emph{C. cylindracea} were approximately in-between the values of mixed
and monospecific \emph{C. cylindracea}. To assess seasonal changes in
the proportion of shared OTUs and communities the Jaccard's and
Bray-Curtis Similarity Coefficients were calculated between consecutive
sampling points (\autoref{shared}). Both coefficients showed similar
trends. Temporal proportional changes were more pronounced for ambient
seawater than for \emph{C. nodosa} and especially \emph{C. cylindracea}
associated communities (\autoref{shared}). In addition, only 0.4 -- 1.0
\si{\percent} of OTUs from each surface associated community were
present at all seasons. These persistent OTUs constituted a high
proportion of total sequences (40.2 -- 52.1 \si{\percent}). To further
disentangle the environmental and seasonal community dissimilarity a
Principal Coordinates Analysis (PCoA) based on Bray-Curtis distances and
OTU abundances was applied. A clear separation between ambient seawater
and surface associated communities was found (\autoref{pcoa}). In
addition, a separation of epiphytic bacterial and archaeal communities
based on host species was detected. This separation was further
supported by ANOSIM (R = 0.96, \emph{p} \textless{} 0.001). Seasonal
changes of \emph{C. nodosa} associated communities indicated a
separation between spring, summer and autumn/winter samples (ANOSIM, R =
0.54, \emph{p} \textless{} 0.01). For \emph{C. cylindracea} associated
communities a separation between summer and autumn/winter/spring samples
was observed that was, however, not as strong as for \emph{C. nodosa}
associated communities (ANOSIM, R = 0.30, \emph{p} \textless{} 0.05)
(\autoref{pcoa}).

The taxonomic composition of both, macrophyte associated and ambient
seawater communities was dominated by bacterial (99.1 ± 2.1
\si{\percent}) over archaeal sequences (0.9 ± 2.1 \si{\percent})
(\autoref{community}). Higher relative abundances of chloroplast related
sequences were only observed in surface associated communities, with
higher values in autumn/winter (37.2 ± 11.2 \si{\percent}) than in
spring/summer (20.9 ± 9.7 \si{\percent}) (\autoref{supp-chloroplast}).
Generally, at higher taxonomic ranks (phylum-class), epiphytic and
ambient seawater microbial communities were composed of similar
bacterial taxa. Ambient seawater communities were mainly comprised of
\emph{Actinobacteriota}, \emph{Bacteroidota}, \emph{Cyanobacteria},
\emph{Alphaproteobacteria}, \emph{Gammaproteobacteria} and
\emph{Verrucomicrobiota}. Communities associated with \emph{C. nodosa}
consisted additionally of \emph{Planctomycetota} contributing more in
summer 2018 than in other seasons. In addition, communities from mixed
and monospecific \emph{C. cylindracea} were similar and characterized by
the same groups as ambient seawater and \emph{C. nodosa} communities
with the addition of \emph{Desulfobacterota} (\autoref{community}).
Larger differences between environments and host species were observed
at lower taxonomic ranks (Figs. \ref{cyano} -- \ref{desulfo}).

\emph{Cyanobacteria} related sequences comprised, on average, 5.5 ± 4.4
\si{\percent} of total sequences (\autoref{cyano}). Higher proportions
were found for \emph{C. nodosa} (16.4 ± 5.3 \si{\percent}) and \emph{C.
cylindracea} mixed (7.7 ± 3.9 \si{\percent}) and monospecific (7.8 ± 2.4
\si{\percent}) associated communities in autumn and for ambient seawater
communities in winter (8.8 ± 7.5 \si{\percent}). Large taxonomic
differences between surface associated and ambient seawater
cyanobacterial communities were observed. Ambient seawater communities
were mainly comprised of \emph{Cyanobium} and \emph{Synechococcus},
while surface associated communities were comprised of
\emph{Pleurocapsa} and sequences within the class \emph{Cyanobacteriia}
that could not be further classified (no relative \emph{Cyanobacteriia})
(\autoref{cyano}). In addition, seasonal changes in surface associated
communities were observed in \emph{Pleurocapsa} and no relative
\emph{Cyanobacteriia} comprising larger proportions in autumn and winter
and \emph{Acrophormium}, \emph{Phormidesmis} and sequences without known
relatives within the \emph{Nodosilineaceae} (no relative
\emph{Nodosilineaceae}) in spring and summer (\autoref{cyano}).

Sequences classified as \emph{Bacteroidota} comprised, on average, 19.2
± 5.5 \si{\percent} of all sequences (\autoref{bactero}). Similarly to
\emph{Cyanobacteria}, large differences in the taxonomic composition
between ambient seawater and surface associated communities were found
(\autoref{bactero}). The ambient seawater community was characterized by
the NS4 and NS5 marine groups, uncultured \emph{Cryomorphaceae},
uncultured \emph{Flavobacteriaceae}, NS11-12 marine group,
\emph{Balneola}, uncultured \emph{Balneolaceae} and \emph{Formosa}. In
contrast, in surface associated communities \emph{Lewinella},
\emph{Portibacter}, \emph{Rubidimonas}, sequences without known
relatives within the \emph{Saprospiraceae} (no relative
\emph{Saprospiraceae}), uncultured \emph{Saprospiraceae}, (sequences
without known relatives within the \emph{Flavobacteriaceae} (no relative
\emph{Flavobacteriaceae})and uncultured \emph{Rhodothermaceae} were
found. Some groups showed minor seasonal changes such as no relative
\emph{Flavobacteriaceae} whose sequences were more abundant from
November 2017 until June 2018. In contrast, uncultured
\emph{Rhodothermaceae} showed higher proportions from June 2018 until
the end of the study period. Surface associated \emph{Bacteroidota}
communities were very diverse as observed in the high proportion of taxa
clustering as other \emph{Bacteroidota} (\autoref{bactero}).

On average, \emph{Alphaproteobacteria} were in comparison to the other
high rank taxa the largest taxonomic group, comprising 29.2 ± 12.0
\si{\percent} of all sequences (\autoref{alpha}). In accordance to the
above described taxa, large differences between ambient seawater and
surface associated communities were observed. Ambient seawater
communities were composed mainly of the SAR11 clade, AEGEAN-169 marine
group, SAR116 clade, sequences without known relatives within the
\emph{Rhodobacteraceae} (no relative \emph{Rhodobacteraceae}), HIMB11
and the OCS116 clade, while surface associated communities were composed
mainly of no relative \emph{Rhodobacteraceae} and to a lesser degree of
\emph{Pseudoahrensia}, \emph{Amylibacter} and sequences without known
relatives within the \emph{Alphaproteobacteria} (no relative
\emph{Alphaproteobacteria}) and \emph{Hyphomonadaceae} (no relative
\emph{Hyphomonadaceae}). Representatives of no relative
\emph{Rhodobacteraceae} comprised on average 40.6 ± 23.2 \si{\percent}
of all alphaproteobacterial sequences in the epiphytic community
(\autoref{alpha}). In addition, \emph{Amylibacter} was detected mainly
in \emph{C. nodosa} from November 2017 until March 2018.

Sequences related to \emph{Gammaproteobacteria} comprised on average
18.6 ± 3.9 \si{\percent} of all sequences (\autoref{gamma}). Similar to
above mentioned taxa, large taxonomic differences between ambient
seawater and surface associated communities were found. Ambient seawater
communities were mainly comprised of the OM60 (NOR5) clade,
\emph{Litoricola}, \emph{Acinetobacter} and the SAR86 clade, while
epiphytic communities were mainly composed of sequences without known
relatives within the \emph{Gammaproteobacteria} (no relative
\emph{Gammaproteobacteria}) and \emph{Granulosicoccus}. Beside these two
groups specific to all three epiphytic communities, \emph{C. nodosa} was
characterized by \emph{Arenicella}, \emph{Methylotenera} and sequences
without known relatives within the \emph{Burkholderiales} (no relative
\emph{Burkholderiales}), while \emph{Thioploca}, \emph{Reinekea} and
sequences without known relatives within \emph{Cellvibrionaceae} (no
relative \emph{Cellvibrionaceae}) were more specific to both mixed and
monospecific \emph{C. cylindracea}. In addition, \emph{Arenicella} was
more pronounced in November and December 2017, while no relative
\emph{Burkholderiales} and \emph{Methylotenera} were characteristic for
the period from March until May 2018. For the \emph{C. cylindracea}
specific taxa no relative \emph{Cellvibrionaceae} and \emph{Reinekea}
showed seasonality and were characteristic for samples originating from
June to October 2018. In addition, similar to \emph{Bacteroidota}, a
large proportion of the surface associated community was grouped as
other \emph{Gammaproteobacteria} indicating high diversity within this
group (\autoref{gamma}).

\emph{Desulfobacterota} were specific for \emph{C. cylindracea}. In the
mixed and monospecific \emph{C. cylindracea} communities the proportion
of \emph{Desulfobacterota} was 25.7 ± 11.2 \si{\percent} and 24.0 ± 4.3
\si{\percent}, respectively (\autoref{desulfo}). In contrast, in ambient
seawater and \emph{C. nodosa} communities the contribution of
\emph{Desulfobacterota} was only 0.1 ± 0.08 \si{\percent} and 1.0 ± 0.7
\si{\percent}, respectively. In \emph{C. cylindracea} the community
consisted mainly of \emph{Desulfatitalea}, \emph{Desulfobulbus},
\emph{Desulfopila}, \emph{Desulforhopalus}, \emph{Desulfotalea},
SEEP-SRB4, uncultured \emph{Desulfocapsaceae} and sequences without
known relatives within the \emph{Desulfobacteraceae} (no relative
\emph{Desulfobacteraceae}), \emph{Desulfobulbaceae} (no relative
\emph{Desulfobulbaceae}) and \emph{Desulfocapsaceae} (no relative
\emph{Desulfocapsaceae}) (\autoref{desulfo}).

\newpage

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

Surfaces of marine macrophytes harbor biofilms consisting of diverse
microbial taxa (Egan et al., 2013; Tarquinio et al., 2019). No standard
protocol has been developed to study these macophyte-associated microbes
(Ugarelli et al., 2019). Different procedures for removal of microbial
cells from host surfaces are described, such as host tissue shaking
(Nõges et al., 2010), scraping (Uku et al., 2007), swabbing (Mancuso et
al., 2016) and ultrasonication (Cai et al., 2014). All these methods
result in different removal efficiencies but none was enabling a
complete removal of attached microbial cells based on our experience. In
the present study, we applied a removal protocol (Korlević et al.,
submitted) based on direct cellular lysis (Burke et al., 2009). The
application of a direct lysis procedure coupled with a high sampling
frequency and Illumina amplicon sequencing has enabled us to described
in detail the bacterial and archaeal communities associated with the
surfaces of two marine macrophytes, \emph{C. nodosa} and \emph{C.
cylindracea}.

In the present study, highest richness was observed for \emph{C.
cylindracea} (mixed and monospecific) followed by \emph{C. nodosa} and
lowest richness was found in ambient seawater microbial communities.
Higher richness of microbial communities associated with seagrasses than
in ambient seawater was described earlier and could be attributed to a
larger set of inhabitable microniches existing on macrophyte surfaces
than in the ambient seawater (Ugarelli et al., 2019). The highest
richness observed for \emph{C. cylindracea} might be partly due to its
contact with the sediment. The stolon of \emph{C. cylindracea} is
attached to the sediment surface with rhizoids and thus, the stolon and
rhizoids are in a direct contact with the sediment. In addition,
seasonal differences in richness observed for surface attached
communities indicated a slightly higher richness in spring and summer.
This pattern could be explained by a higher macrophyte growth in these
two seasons than in autumn and winter (M. Najdek, personal
communication; Zavodnik et al., 1998; Ruitton et al., 2005). During
their main growth season in spring and summer macrophytes exhibit a more
dynamic chemical interaction with the surface community probably causing
an increase in the number of inhabitable microniches (Borges and
Champenois, 2015; Rickert et al., 2016).

We observed a strong differentiation between the surface attached and
ambient seawater communities at the level of OTUs, in agreement with
most published studies (Burke et al., 2011b; Michelou et al., 2013;
Mancuso et al., 2016; Roth-Schulze et al., 2016; Crump et al., 2018;
Ugarelli et al., 2019). This indicates that marine macrophytes are a
selecting factor from the pool of microbial taxa present in the ambient
seawater, modifying the microbial community once the macrophyte
associated microbial biofilm develops (Salaün et al., 2012; Michelou et
al., 2013). In contrast, Fahimipour et al. (2017) report in a global
study of \emph{Zostera marina}, similarities between the microbial
community developed on leaves and in the ambient seawater. The
discrepancy between our data and the study of Fahimipour et al. (2017)
could be explained by different seagrass species, methodological
variations or biogeographic trends as Fahimipour et al. (2017) analysed
samples from different locations throughout the northern hemisphere. It
is possible that the microbial communities in ambient seawater and on
leaves from the same location are differing but are still more similar
to each other when compared to other sampling locations. Indeed, it was
found that prokaryotic communities vary substantially between different
sampling sites (Bengtsson et al., 2017). When the taxonomic composition
at high ranks was analysed no such strong differentiation was noticed.
Phyla and classes such as \emph{Actinobacteriota}, \emph{Bacteroidota},
\emph{Cyanobacteria}, \emph{Alphaproteobacteria},
\emph{Gammaproteobacteria} and \emph{Verrucomicrobiota} were found in
both ambient seawater as well as macrophyte associated, in agreement
with previous studies (Burke et al., 2011b; Egan et al., 2013; Michelou
et al., 2013). In contrast, when low taxonomic ranks were analysed
(i.e., family and genus) a strong differentiation was observed (Figs.
\ref{cyano} -- \ref{desulfo}). A similar differentiation at lower
microbial taxonomic ranks between ambient seawater and macrophytes was
described for other macrophyte species as well (Egan et al., 2013;
Michelou et al., 2013; Ugarelli et al., 2019).

Beside differences between ambient seawater and surface associated
microbial communities, it is unclear whether the prokaryotic epiphytic
community is host-specific or whether there are generalist taxa
characteristic to all or many macrophytes (Egan et al., 2013). Similar
to previously described differences between microbial communities in the
ambient seawater and on macrophytes, at high taxonomic ranks no major
difference between the microbial communities associated with different
hosts was observed. The only high rank phylum that was differing between
\emph{C. nodosa} and \emph{C. cylindracea} was \emph{Desulfobacterota},
with more abundant sequences in the \emph{C. cylindracea} associated
community. As already mentioned, the rhizoids and part of the stolon are
in contact with the sediment. Thus \emph{Desulfobacterota} are probably
a part of the epiphytic community that was in contact with the sediment.
Similar high rank taxa found in this study were described to be specific
for other species of macrophytes (Burke et al., 2011b; Lachnit et al.,
2011; Mancuso et al., 2016; Bengtsson et al., 2017). In contrast to high
taxonomic ranks, a substantial differentiation between host specific
communities was found supporting the notion that macrophyte associated
microbial communities might be host-specific. Host-specificity was also
observed for other species of macroalgae and seagrasses (Lachnit et al.,
2011; Roth-Schulze et al., 2016; Morrissey et al., 2019; Ugarelli et
al., 2019). Taken together, at high taxonomic ranks a core set of taxa
could be described that is characteristic for all or many macrophytes,
while at low taxonomic ranks a community specific to host species was
identified (Figs. \ref{pcoa} and \ref{community}) (Egan et al., 2013).

Seasonal changes in richness in the epiphytic community were substantial
as indicated by the proportion of OTUs (\(\leq\) 1.0 \si{\percent})
present at every sampling date. These persistent OTUs, however, were
accounting for a high proportion of sequences (\(\geq\) 40.2
\si{\percent}) (\autoref{shared}). A very similar proportion of
persistent OTUs was reported in high-frequency sampling studies
describing seasonal changes in picoplankton (Gilbert et al., 2009,
2012). In comparison to the seawater community, a lower degree of
seasonal shifts was observed for the macrophyte surface associated
communities. It appears that microniches at the surfaces of macrophytes
are providing more stable conditions than the ambient seawater. At the
level of OTUs seasonal changes of \emph{C. nodosa} and \emph{C.
cylindracea} associated communities were identified that could be linked
to the growth cycle of the seagrass and macroalgae (M. Najdek, personal
communication). \emph{C. nodosa} was characterized by a spring community
during maximum seagrass proliferation, a summer community during the
highest standing stock of \emph{C.nodosa} and an autumn/winter community
during the decay of seagrass biomass. In contrast, \emph{C. cylindracea}
started to proliferate in late spring and was characterized only by a
summer community during high growth rates and by an autumn/winter/spring
community when the biomass was at the peak and decaying thereafter.
Similar seasonal changes in the epiphytic community were also described
for other macroalgae (Tujula et al., 2010; Lachnit et al., 2011). Higher
seasonal stability of \emph{C. cylindracea} surface communities than in
\emph{C. nodosa} was also observed in the higher proportion of shared
communities between two consecutive sampling dates in \emph{C.
cylindracea}.

Chloroplast sequence abundances were higher in autumn/winter than in
spring/summer. This pattern is not surprising as seagrasses harbor more
algal epiphytes during autumn/winter than in spring/summer (Reyes and
Sansón, 2001). Furthermore, we used an adapted DNA isolation protocol
that is known to partially co-extract DNA from planktonic eukaryotes
(Korlević et al., 2015). Strong seasonal fluctuations of high rank
epiphytic taxa were not observed, with the exception of
\emph{Cyanobacteria}. Cyanobacterial sequences were more pronounced in
November and December than in spring and summer. In the months of high
cyanobacterial sequence abundances the majority of sequences from this
group were classified as \emph{Pleurocapsa}, a group known to colonized
different living and non-living surfaces (Burns et al., 2004; Longford
et al., 2007; Mobberley et al., 2012; Reisser et al., 2014). It is
possible that during periods of low metabolic activity there is a
reduced interaction and selection of the epiphytic community by the
seagrass, causing leaves to become a suitable surface for non-specific
colonizers (Zavodnik et al., 1998). \emph{Pleurocapsa} was replaced in
spring and summer by \emph{Acrophormium}, \emph{Phormidesmis} and
sequences without known relatives within the \emph{Nodosilineaceae}. A
study of coastal microbial mats found also a higher proportion of
\emph{Nodosilineaceae} sequences in summer, while \emph{Phormidesmis}
sequences were at their peak in autumn (Cardoso et al., 2019). Other
high rank taxa did not exhibit strong successional patterns. In every
analysed group, with the exception of \emph{Desulfobacterota}, taxa
present throughout the year in similar proportions and season specific
taxa could be identified (Figs. \ref{bactero} and \ref{desulfo}). Within
\emph{Bacteroidota} different groups within the family
\emph{Saprospiraceae} (e.g. \emph{Lewinella}, \emph{Portibacter} and
\emph{Rubidimonas}) were detected across all seasons. Members of this
family are often found in association with macrophytes and it is
suggested that they are involved in the hydrolysis and utilization of
complex carbon sources (Burke et al., 2011b; McIlroy and Nielsen, 2014;
Crump et al., 2018). In contrast, the families \emph{Flavobacteriaceae}
and \emph{Rhodothermaceae} showed seasonal patterns, with
\emph{Flavobacteriaceae} being more pronounced from November to June and
\emph{Rhodothermaceae} from June to October (\autoref{bactero}). Within
\emph{Alphaproteobacteria} the family \emph{Rhodobacteraceae} comprised
the majority of sequences throughout the year (\autoref{alpha}). This
metabolically versatile family is often associated with macrophyte
surfaces and usually is one of the most abundant groups (Burke et al.,
2011b; Michelou et al., 2013; Pujalte et al., 2014; Mancuso et al.,
2016). In addition, \emph{Hyphomonadaceae} were found in all samples.
Interestingly, some of the species within this group contain stalks on
their cells, which can be used to attach to the macrophyte surface
(Weidner et al., 2000; Abraham and Rohde, 2014). Within the
\emph{Gammaproteobacteria}, sequences without known representatives were
the most pronounced group present throughout the year (\autoref{gamma}).
In addition, \emph{Granulosicoccus} was also found in almost all
samples. \emph{Gammaproteobacteria} are often a major constituent of
macrophyte epiphytic communities (Burke et al., 2011b; Michelou et al.,
2013; Crump et al., 2018). Beside these two groups, other less abundant,
taxa showed seasonal and host-specific patterns. For example, \emph{C.
cylindracea} harbored \emph{Thioploca}, a known sulfur sediment bacteria
and \emph{Cellvibrionaceae}, a family with cultured members known as
polysaccharide degraders (Jørgensen and Gallardo, 1999; Xie et al.,
2017). \emph{Desulfobacterota} were found only associated with \emph{C.
cylindracea} and no group within this phylum showed seasonal patterns
(\autoref{desulfo}). The presence of this phylum only on \emph{C.
cylindracea} is to be expected as part of the epiphytic community is in
direct contact with the sediment. The \emph{Desulfobacterota} community
was dominated by \emph{Desulfatitalea} and no relative
\emph{Desulfocapsaceae}, known sulphate sediment groups (Kuever, 2014;
Higashioka et al., 2015).

In temperate zones, marine macrophytes are exhibiting growth cycles, so
it is not surprising that the associated epiphytic microbial community
is undergoing partial seasonal changes. In the present study, we could
identify in every analysed high rank taxa phylogenetic groups present
throughout the year, comprising most of the sequences and a lower
proportion of taxa showing seasonal patterns connected to the macrophyte
growth cycle (Figs. \ref{community} and \ref{desulfo}). Studies focusing
on functional comparisons between communities associated with different
hosts showed that the majority of functions could be found in every
community, indicating functional redundancy (Roth-Schulze et al., 2016).
This difference between phylogenetic varaibility and functional
stability was explained by the lottery hypothesis assuming an initial
random colonization step performed by a set of functionally equivalent
taxonomic groups (Burke et al., 2011a; Roth-Schulze et al., 2016). It is
possible that functional redundancy is a characteristic of high
abundance taxa detected to be present throughout the year, while
seasonal and/or host-specific functions are an attribute of taxa
displaying successional patterns. Further studies connecting taxonomy
with functional properties will be required to elucidate the degree of
functional redundancy or specificity in epiphytic microbial communities.

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

This work was founded by the Croatian Science Foundation through the
MICRO-SEAGRASS project (project number IP-2016-06-7118). ZZ and GJH were
supported by the Austrian Science Fund (FWF) project ARTEMIS (project
number P28781-B21). We would like to thank Margareta Buterer for
technical support, Paolo Paliaga for help during sampling and the
University Computing Center of the University of Zagreb for access to
the computer cluster Isabella.

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\interlinepenalty=10000 \setlength{\emergencystretch}{3.5em}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abraham2014}{}%
Abraham, W. R., and Rohde, M. (2014). ``The family
\emph{Hyphomonadaceae},'' in \emph{The Prokaryotes: Alphaproteobacteria
and Betaproteobacteria}, eds. E. Rosenberg, E. F. DeLong, S. Lory, E.
Stackebrandt, and F. Thompson (Berlin, Heidelberg: Springer-Verlag),
283--299.
doi:\href{https://doi.org/10.1007/978-3-642-30197-1_260}{10.1007/978-3-642-30197-1\_260}.

\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J. J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019). \emph{rmarkdown: Dynamic documents for R}.

\leavevmode\hypertarget{ref-Apprill2015}{}%
Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015). Minor
revision to V4 region SSU rRNA 806R gene primer greatly increases
detection of SAR11 bacterioplankton. \emph{Aquat. Microb. Ecol.} 75,
129--137. doi:\href{https://doi.org/10.3354/ame01753}{10.3354/ame01753}.

\leavevmode\hypertarget{ref-Armstrong2000}{}%
Armstrong, E., Rogerson, A., and Leftley, J. (2000). The abundance of
heterotrophic protists associated with intertidal seaweeds.
\emph{Estuar. Coast. Shelf Sci.} 50, 415--424.
doi:\href{https://doi.org/10.1006/ECSS.1999.0577}{10.1006/ECSS.1999.0577}.

\leavevmode\hypertarget{ref-Bengtsson2017}{}%
Bengtsson, M. M., Bühler, A., Brauer, A., Dahlke, S., Schubert, H., and
Blindow, I. (2017). Eelgrass leaf surface microbiomes are locally
variable and highly correlated with epibiotic eukaryotes. \emph{Front.
Microbiol.} 8, 1312.
doi:\href{https://doi.org/10.3389/fmicb.2017.01312}{10.3389/fmicb.2017.01312}.

\leavevmode\hypertarget{ref-Bengtsson2010}{}%
Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010). Seasonal dynamics of
bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat.
Microb. Ecol.} 60, 71--83.
doi:\href{https://doi.org/10.3354/ame01409}{10.3354/ame01409}.

\leavevmode\hypertarget{ref-Borges2015}{}%
Borges, A. V., and Champenois, W. (2015). Seasonal and spatial
variability of dimethylsulfoniopropionate (DMSP) in the Mediterranean
seagrass \emph{Posidonia oceanica}. \emph{Aquat. Bot.} 125, 72--79.
doi:\href{https://doi.org/10.1016/j.aquabot.2015.05.008}{10.1016/j.aquabot.2015.05.008}.

\leavevmode\hypertarget{ref-Burke2009}{}%
Burke, C., Kjelleberg, S., and Thomas, T. (2009). Selective extraction
of bacterial DNA from the surfaces of macroalgae. \emph{Appl. Environ.
Microbiol.} 75, 252--256.
doi:\href{https://doi.org/10.1128/AEM.01630-08}{10.1128/AEM.01630-08}.

\leavevmode\hypertarget{ref-Burke2011}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011a). Bacterial community assembly based on functional genes rather
than species. \emph{Proc. Natl. Acad. Sci. U. S. A.} 108, 14288--14293.
doi:\href{https://doi.org/10.1073/pnas.1101591108}{10.1073/pnas.1101591108}.

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
(2011b). Composition, uniqueness and variability of the epiphytic
bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
J.} 5, 590--600.
doi:\href{https://doi.org/10.1038/ismej.2010.164}{10.1038/ismej.2010.164}.

\leavevmode\hypertarget{ref-Burns2004}{}%
Burns, B. P., Goh, F., Allen, M., and Neilan, B. A. (2004). Microbial
diversity of extant stromatolites in the hypersaline marine environment
of Shark Bay, Australia. \emph{Environ. Microbiol.} 6, 1096--1101.
doi:\href{https://doi.org/10.1111/j.1462-2920.2004.00651.x}{10.1111/j.1462-2920.2004.00651.x}.

\leavevmode\hypertarget{ref-Cai2014}{}%
Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., et al. (2014).
An ultrasonic method for separation of epiphytic microbes from
freshwater submerged macrophytes. \emph{J. Basic Microbiol.} 54,
758--761.
doi:\href{https://doi.org/10.1002/jobm.201300041}{10.1002/jobm.201300041}.

\leavevmode\hypertarget{ref-Caporaso2012}{}%
Caporaso, J. G., Lauber, C. L., Walters, W. A., Berg-Lyons, D., Huntley,
J., Fierer, N., et al. (2012). Ultra-high-throughput microbial community
analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J.} 6,
1621--1624.
doi:\href{https://doi.org/10.1038/ismej.2012.8}{10.1038/ismej.2012.8}.

\leavevmode\hypertarget{ref-Cardoso2019}{}%
Cardoso, D. C., Cretoiu, M. S., Stal, L. J., and Bolhuis, H. (2019).
Seasonal development of a coastal microbial mat. \emph{Sci. Rep.} 9,
9035.
doi:\href{https://doi.org/10.1038/s41598-019-45490-8}{10.1038/s41598-019-45490-8}.

\leavevmode\hypertarget{ref-Crump2008}{}%
Crump, B. C., and Koch, E. W. (2008). Attached bacterial populations
shared by four species of aquatic angiosperms. \emph{Appl. Environ.
Microbiol.} 74, 5948--5957.
doi:\href{https://doi.org/10.1128/AEM.00952-08}{10.1128/AEM.00952-08}.

\leavevmode\hypertarget{ref-Crump2018}{}%
Crump, B. C., Wojahn, J. M., Tomas, F., and Mueller, R. S. (2018).
Metatranscriptomics and amplicon sequencing reveal mutualisms in
seagrass microbiomes. \emph{Front. Microbiol.} 9, 388.
doi:\href{https://doi.org/10.3389/fmicb.2018.00388}{10.3389/fmicb.2018.00388}.

\leavevmode\hypertarget{ref-Egan2013}{}%
Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
Thomas, T. (2013). The seaweed holobiont: Understanding seaweed-bacteria
interactions. \emph{FEMS Microbiol. Rev.} 37, 462--476.
doi:\href{https://doi.org/10.1111/1574-6976.12011}{10.1111/1574-6976.12011}.

\leavevmode\hypertarget{ref-Fahimipour2017}{}%
Fahimipour, A. K., Kardish, M. R., Lang, J. M., Green, J. L., Eisen, J.
A., and Stachowicz, J. J. (2017). Global-scale structure of the eelgrass
microbiome. \emph{Appl. Environ. Microbiol.} 83, e03391--16.
doi:\href{https://doi.org/10.1128/AEM.03391-16}{10.1128/AEM.03391-16}.

\leavevmode\hypertarget{ref-Gilbert2009}{}%
Gilbert, J. A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth,
T., et al. (2009). The seasonal structure of microbial communities in
the Western English Channel. \emph{Environ. Microbiol.} 11, 3132--3139.
doi:\href{https://doi.org/10.1111/j.1462-2920.2009.02017.x}{10.1111/j.1462-2920.2009.02017.x}.

\leavevmode\hypertarget{ref-Gilbert2012}{}%
Gilbert, J. A., Steele, J. A., Caporaso, J. G., Steinbrück, L., Reeder,
J., Temperton, B., et al. (2012). Defining seasonal marine microbial
community dynamics. \emph{ISME J.} 6, 298--308.
doi:\href{https://doi.org/10.1038/ismej.2011.107}{10.1038/ismej.2011.107}.

\leavevmode\hypertarget{ref-Higashioka2015}{}%
Higashioka, Y., Kojima, H., Watanabe, T., and Fukui, M. (2015). Draft
genome sequence of \emph{Desulfatitalea tepidiphila}
S28bF\textsuperscript{T}. \emph{Genome Announc.} 3, e01326--15.
doi:\href{https://doi.org/10.1128/genomeA.01326-15}{10.1128/genomeA.01326-15}.

\leavevmode\hypertarget{ref-Hollants2013}{}%
Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013). What
we can learn from sushi: A review on seaweed-bacterial associations.
\emph{FEMS Microbiol. Ecol.} 83.
doi:\href{https://doi.org/10.1111/j.1574-6941.2012.01446.x}{10.1111/j.1574-6941.2012.01446.x}.

\leavevmode\hypertarget{ref-Jost2006}{}%
Jost, L. (2006). Entropy and diversity. \emph{Oikos} 113, 363--375.
doi:\href{https://doi.org/10.1111/j.2006.0030-1299.14714.x}{10.1111/j.2006.0030-1299.14714.x}.

\leavevmode\hypertarget{ref-Jorgensen1999}{}%
Jørgensen, B. B., and Gallardo, V. A. (1999). \emph{Thioploca} spp.:
filamentous sulfur bacteria with nitrate vacuoles. \emph{FEMS Microbiol.
Ecol.} 28, 301--313.
doi:\href{https://doi.org/10.1111/j.1574-6941.1999.tb00585.x}{10.1111/j.1574-6941.1999.tb00585.x}.

\leavevmode\hypertarget{ref-Korlevic}{}%
Korlević, M., Markovski, M., Zhao, Z., Herndl, G. J., and Najdek, M.
Selective DNA and protein isolation from marine macrophyte surfaces.

\leavevmode\hypertarget{ref-Korlevic2015}{}%
Korlević, M., Pop Ristova, P., Garić, R., Amann, R., and Orlić, S.
(2015). Bacterial diversity in the South Adriatic Sea during a strong,
deep winter convection year. \emph{Appl. Environ. Microbiol.} 81,
1715--1726.
doi:\href{https://doi.org/10.1128/AEM.03410-14}{10.1128/AEM.03410-14}.

\leavevmode\hypertarget{ref-Kozich2013}{}%
Kozich, J. J., Westcott, S. L., Baxter, N. T., Highlander, S. K., and
Schloss, P. D. (2013). Development of a dual-index sequencing strategy
and curation pipeline for analyzing amplicon sequence data on the MiSeq
Illumina sequencing platform. \emph{Appl. Environ. Microbiol.} 79,
5112--5120.
doi:\href{https://doi.org/10.1128/AEM.01043-13}{10.1128/AEM.01043-13}.

\leavevmode\hypertarget{ref-Kuever2014}{}%
Kuever, J. (2014). ``The family \emph{Desulfobulbaceae},'' in \emph{The
Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}, eds. E.
Rosenberg, E. F. DeLong, S. Lory, E. Stackebrandt, and F. Thompson
(Berlin, Heidelberg: Springer-Verlag), 75--86.
doi:\href{https://doi.org/10.1007/978-3-642-39044-9_267}{10.1007/978-3-642-39044-9\_267}.

\leavevmode\hypertarget{ref-Lachnit2009}{}%
Lachnit, T., Blümel, M., Imhoff, J. F., and Wahl, M. (2009). Specific
epibacterial communities on macroalgae: Phylogeny matters more than
habitat. \emph{Aquat. Biol.} 5, 181--186.
doi:\href{https://doi.org/10.3354/ab00149}{10.3354/ab00149}.

\leavevmode\hypertarget{ref-Lachnit2011}{}%
Lachnit, T., Meske, D., Wahl, M., Harder, T., and Schmitz, R. (2011).
Epibacterial community patterns on marine macroalgae are host-specific
but temporally variable. \emph{Environ. Microbiol.} 13, 655--665.
doi:\href{https://doi.org/10.1111/j.1462-2920.2010.02371.x}{10.1111/j.1462-2920.2010.02371.x}.

\leavevmode\hypertarget{ref-Longford2007}{}%
Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
Kjelleberg, S., et al. (2007). Comparisons of diversity of bacterial
communities associated with three sessile marine eukaryotes.
\emph{Aquat. Microb. Ecol.} 48, 217--229.
doi:\href{https://doi.org/10.3354/ame048217}{10.3354/ame048217}.

\leavevmode\hypertarget{ref-Mancuso2016}{}%
Mancuso, F. P., D'Hondt, S., Willems, A., Airoldi, L., and De Clerck, O.
(2016). Diversity and temporal dynamics of the epiphytic bacterial
communities associated with the canopy-forming seaweed \emph{Cystoseira
compressa} (Esper) Gerloff and Nizamuddin. \emph{Front Microbiol} 7,
476.
doi:\href{https://doi.org/10.3389/fmicb.2016.00476}{10.3389/fmicb.2016.00476}.

\leavevmode\hypertarget{ref-Margulis1991}{}%
Margulis, L. (1991). ``Symbiogenesis and symbionticism,'' in
\emph{Symbiosis as a Source of Evolutionary Innovation: Speciation and
Morphogenesis}, eds. L. Margulis and R. Fester (Cambridge,
Massachusetts: The MIT Press), 1--14.

\leavevmode\hypertarget{ref-Massana1997}{}%
Massana, R., Murray, A. E., Preston, C. M., and DeLong, E. F. (1997).
Vertical distribution and phylogenetic characterization of marine
planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl.
Environ. Microbiol.} 63, 50--56.

\leavevmode\hypertarget{ref-McIlroy2014}{}%
McIlroy, S. J., and Nielsen, P. H. (2014). ``The family
\emph{Saprospiraceae},'' in \emph{The Prokaryotes: Other Major Lineages
of Bacteria and the Archaea}, eds. E. Rosenberg, E. F. DeLong, S. Lory,
E. Stackebrandt, and F. Thompson (Berlin, Heidelberg: Springer-Verlag),
863--889.
doi:\href{https://doi.org/10.1007/978-3-642-38954-2_138}{10.1007/978-3-642-38954-2\_138}.

\leavevmode\hypertarget{ref-Michelou2013}{}%
Michelou, V. K., Caporaso, J. G., Knight, R., and Palumbi, S. R. (2013).
The ecology of microbial communities associated with \emph{Macrocystis
pyrifera}. \emph{PloS One} 8, e67480.
doi:\href{https://doi.org/10.1371/journal.pone.0067480}{10.1371/journal.pone.0067480}.

\leavevmode\hypertarget{ref-Miranda2013}{}%
Miranda, L. N., Hutchison, K., Grossman, A. R., and Brawley, S. H.
(2013). Diversity and abundance of the bacterial community of the red
macroalga \emph{Porphyra umbilicalis}: Did bacterial farmers produce
macroalgae? \emph{PloS One} 8, e58269.
doi:\href{https://doi.org/10.1371/journal.pone.0058269}{10.1371/journal.pone.0058269}.

\leavevmode\hypertarget{ref-Mobberley2012}{}%
Mobberley, J. M., Ortega, M. C., and Foster, J. S. (2012). Comparative
microbial diversity analyses of modern marine thrombolitic mats by
barcoded pyrosequencing. \emph{Environ. Microbiol.} 14, 82--100.
doi:\href{https://doi.org/10.1111/j.1462-2920.2011.02509.x}{10.1111/j.1462-2920.2011.02509.x}.

\leavevmode\hypertarget{ref-Morrissey2019}{}%
Morrissey, K. L., Çavas, L., Willems, A., and De Clerck, O. (2019).
Disentangling the influence of environment, host specificity and thallus
differentiation on bacterial communities in siphonous green seaweeds.
\emph{Front. Microbiol.} 10, 717.
doi:\href{https://doi.org/10.3389/fmicb.2019.00717}{10.3389/fmicb.2019.00717}.

\leavevmode\hypertarget{ref-Neuwirth2014}{}%
Neuwirth, E. (2014). \emph{RColorBrewer: ColorBrewer palettes}.

\leavevmode\hypertarget{ref-Noges2010}{}%
Nõges, T., Luup, H., and Feldmann, T. (2010). Primary production of
aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
Võrtsjärv) in Estonia. \emph{Aquat. Ecol.} 44, 83--92.
doi:\href{https://doi.org/10.1007/s10452-009-9249-4}{10.1007/s10452-009-9249-4}.

\leavevmode\hypertarget{ref-Oksanen2019}{}%
Oksanen, J., Blanchet, F. G., Friendly, M., Kindt, R., Legendre, P.,
McGlinn, D., et al. (2019). \emph{vegan: Community ecology package}.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A. E., Needham, D. M., and Fuhrman, J. A. (2016). Every base
matters: Assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environ. Microbiol.} 18, 1403--1414.
doi:\href{https://doi.org/10.1111/1462-2920.13023}{10.1111/1462-2920.13023}.

\leavevmode\hypertarget{ref-Pujalte2014}{}%
Pujalte, M. J., Lucena, T., Ruvira, M. A., Arahal, D. R., and Macián, M.
C. (2014). ``The family \emph{Rhodobacteraceae},'' in \emph{The
Prokaryotes: Alphaproteobacteria and Betaproteobacteria}, eds. E.
Rosenberg, E. F. DeLong, S. Lory, E. Stackebrandt, and F. Thompson
(Berlin, Heidelberg: Springer-Verlag), 439--512.
doi:\href{https://doi.org/10.1007/978-3-642-30197-1_377}{10.1007/978-3-642-30197-1\_377}.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013). The SILVA ribosomal RNA gene database project: Improved
data processing and web-based tools. \emph{Nucleic Acids Res.} 41,
D590--D596.
doi:\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Reisser2014}{}%
Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D. K. A.,
Thums, M., et al. (2014). Millimeter-sized marine plastics: A new
pelagic habitat for microorganisms and invertebrates. \emph{PloS One} 9,
e100289.
doi:\href{https://doi.org/10.1371/journal.pone.0100289}{10.1371/journal.pone.0100289}.

\leavevmode\hypertarget{ref-Reyes2001}{}%
Reyes, J., and Sansón, M. (2001). Biomass and production of the
epiphytes on the leaves of \emph{Cymodocea nodosa} in the Canary
Islands. \emph{Bot. Mar.} 44, 307--313.
doi:\href{https://doi.org/10.1515/BOT.2001.039}{10.1515/BOT.2001.039}.

\leavevmode\hypertarget{ref-Rickert2016}{}%
Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016).
Seasonal variations in surface metabolite composition of \emph{Fucus
vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PloS
One} 11, e0168196.
doi:\href{https://doi.org/10.1371/journal.pone.0168196}{10.1371/journal.pone.0168196}.

\leavevmode\hypertarget{ref-Roth-Schulze2016}{}%
Roth-Schulze, A. J., Zozaya-Valdés, E., Steinberg, P. D., and Thomas, T.
(2016). Partitioning of functional and taxonomic diversity in
surface-associated microbial communities. \emph{Environ. Microbiol.} 18,
4391--4402.
doi:\href{https://doi.org/10.1111/1462-2920.13325}{10.1111/1462-2920.13325}.

\leavevmode\hypertarget{ref-Ruitton2005}{}%
Ruitton, S., Verlaque, M., and Boudouresque, C. F. (2005). Seasonal
changes of the introduced \emph{Caulerpa racemosa} var.
\emph{cylindracea} (Caulerpales, Chlorophyta) at the northwest limit of
its Mediterranean range. \emph{Aquat. Bot.} 82, 55--70.
doi:\href{https://doi.org/10.1016/j.aquabot.2005.02.008}{10.1016/j.aquabot.2005.02.008}.

\leavevmode\hypertarget{ref-Salaun2012}{}%
Salaün, S., La Barre, S., Santos-Goncalvez, M. D., Potin, P., Haras, D.,
and Bazire, A. (2012). Influence of exudates of the kelp \emph{Laminaria
digitata} on biofilm formation of associated and exogenous bacterial
epiphytes. \emph{Microb. Ecol.} 64, 359--369.
doi:\href{https://doi.org/10.1007/s00248-012-0048-4}{10.1007/s00248-012-0048-4}.

\leavevmode\hypertarget{ref-Schloss2016}{}%
Schloss, P. D., Jenior, M. L., Koumpouras, C. C., Westcott, S. L., and
Highlander, S. K. (2016). Sequencing 16S rRNA gene fragments using the
PacBio SMRT DNA sequencing system. \emph{PeerJ} 4, e1869.
doi:\href{https://doi.org/10.7717/peerj.1869}{10.7717/peerj.1869}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P. D., Westcott, S. L., Ryabin, T., Hall, J. R., Hartmann, M.,
Hollister, E. B., et al. (2009). Introducing mothur: Open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Appl. Environ. Microbiol.} 75,
7537--7541.
doi:\href{https://doi.org/10.1128/AEM.01541-09}{10.1128/AEM.01541-09}.

\leavevmode\hypertarget{ref-Tarquinio2019}{}%
Tarquinio, F., Hyndes, G. A., Laverock, B., Koenders, A., and Säwström,
C. (2019). The seagrass holobiont: Understanding seagrass-bacteria
interactions and their role in seagrass ecosystem functioning.
\emph{FEMS Microbiol. Lett.} 366, fnz057.
doi:\href{https://doi.org/10.1093/femsle/fnz057}{10.1093/femsle/fnz057}.

\leavevmode\hypertarget{ref-Tujula2010}{}%
Tujula, N. A., Crocetti, G. R., Burke, C., Thomas, T., Holmström, C.,
and Kjelleberg, S. (2010). Variability and abundance of the epiphytic
bacterial community associated with a green marine \emph{Ulvacean} alga.
\emph{ISME J.} 4, 301--311.
doi:\href{https://doi.org/10.1038/ismej.2009.107}{10.1038/ismej.2009.107}.

\leavevmode\hypertarget{ref-Ugarelli2019}{}%
Ugarelli, K., Laas, P., and Stingl, U. (2019). The microbial communities
of leaves and roots associated with turtle grass (\emph{Thalassia
testudinum}) and manatee grass (\emph{Syringodium filliforme}) are
distinct from seawater and sediment communities, but are similar between
species and sampling sites. \emph{Microorganisms} 7, 4.
doi:\href{https://doi.org/10.3390/microorganisms7010004}{10.3390/microorganisms7010004}.

\leavevmode\hypertarget{ref-Uku2007}{}%
Uku, J., Björk, M., Bergman, B., and Díez, B. (2007). Characterization
and comparison of prokaryotic epiphytes associated with three East
African seagrasses. \emph{J. Phycol.} 43, 768--779.
doi:\href{https://doi.org/10.1111/j.1529-8817.2007.00371.x}{10.1111/j.1529-8817.2007.00371.x}.

\leavevmode\hypertarget{ref-Weidner2000}{}%
Weidner, S., Arnold, W., Stackebrandt, E., and Pühler, A. (2000).
Phylogenetic analysis of bacterial communities associated with leaves of
the seagrass \emph{Halophila stipulacea} by a culture-independent
small-subunit rRNA gene approach. \emph{Microb. Ecol.} 39, 22--31.

\leavevmode\hypertarget{ref-Wickham2019}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., et al. (2019). Welcome to the tidyverse. \emph{J. Open
Source Softw.} 4, 1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.

\leavevmode\hypertarget{ref-Wilke2018}{}%
Wilke, C. O. (2018). \emph{cowplot: Streamlined plot theme and plot
annotations for 'ggplot2'}.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014). `` knitr: A comprehensive tool for reproducible research
in R,'' in \emph{Implementing Reproducible Computational Research}, eds.
V. Stodden, F. Leisch, and R. D. Peng (New York: Chapman and Hall/CRC),
3--32.

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015). \emph{Dynamic Documents with R and knitr}. 2nd ed. Boca
Raton, Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019). TinyTeX: A lightweight, cross-platform, and
easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
40, 30--32.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R Markdown: The
Definitive Guide}. 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Xie2017}{}%
Xie, Z., Lin, W., and Luo, J. (2017). Comparative phenotype and genome
analysis of \emph{Cellvibrio} sp. PR1, a xylanolytic and agarolytic
bacterium from the Pearl River. \emph{BioMed Res. Int.} 2017, 6304248.
doi:\href{https://doi.org/10.1155/2017/6304248}{10.1155/2017/6304248}.

\leavevmode\hypertarget{ref-Yilmaz2014}{}%
Yilmaz, P., Parfrey, L. W., Yarza, P., Gerken, J., Pruesse, E., Quast,
C., et al. (2014). The SILVA and "All-Species Living Tree Project (LTP)"
taxonomic frameworks. \emph{Nucleic Acids Res.} 42, D643--D648.
doi:\href{https://doi.org/10.1093/nar/gkt1209}{10.1093/nar/gkt1209}.

\leavevmode\hypertarget{ref-Zavodnik1998}{}%
Zavodnik, N., Travizi, A., and de Rosa, S. (1998). Seasonal variations
in the rate of photosynthetic activity and chemical composition of the
seagrass \emph{Cymodocea nodosa} (Ucr.) Asch. \emph{Sci. Mar.} 62,
301--309.
doi:\href{https://doi.org/10.3989/scimar.1998.62n4301}{10.3989/scimar.1998.62n4301}.

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019). \emph{kableExtra: Construct complex table with 'kable'
and pipe syntax}.

\newpage 
\setlength\parindent{0pt}

\hypertarget{figure-legends}{%
\subsection{Figure legends}\label{figure-legends}}

\textbf{\autoref{matrix}.} \nameref{matrix}

\textbf{\autoref{shared}.} \nameref{shared}

\textbf{\autoref{pcoa}.} \nameref{pcoa}

\textbf{\autoref{community}.} \nameref{community}

\textbf{\autoref{cyano}.} \nameref{cyano}

\textbf{\autoref{bactero}.} \nameref{bactero}

\textbf{\autoref{alpha}.} \nameref{alpha}

\textbf{\autoref{gamma}.} \nameref{gamma}

\textbf{\autoref{desulfo}.} \nameref{desulfo}

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{../results/figures/matrix}

}

\caption{Proportion of shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) and shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) between assemblages associated with the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and communities in the ambient seawater.\label{matrix}}\label{fig:unnamed-chunk-1}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/seasonal_shared}

}

\caption{Proportion of shared bacterial and archaeal communities (Bray-Curtis similarity coefficient) and shared bacterial and archaeal OTUs (Jaccard's similarity coefficient) between consecutive sampling dates and from the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\label{shared}}\label{fig:unnamed-chunk-2}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.55\linewidth]{../results/figures/pcoa_figure}

}

\caption{Principal Coordinates Analysis (PCoA) of Bray-Curtis distances based on OTU abundances of bacterial and archaeal communities from the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater. Samples from the same environment or same season are labeled in different colors. The proportion of explained variation by each axis is shown on the corresponding axis in parentheses.\label{pcoa}}\label{fig:unnamed-chunk-3}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/community_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences on the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater. NR -- No Relative (sequences without known relatives within the corresponding group)\label{community}}\label{fig:unnamed-chunk-4}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/cyanobacteria_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) cyanobacterial sequences on the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater. The proportion of cyanobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative (sequences without known relatives within the corresponding group)\label{cyano}}\label{fig:unnamed-chunk-5}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/bacteroidota_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 2 \si{\percent}) sequences within the \textit{Bacteroidota} on the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater. The proportion of sequences classified as \textit{Bacteroidota} in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative (sequences without known relatives within the corresponding group)\label{bactero}}\label{fig:unnamed-chunk-6}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/alphaproteobacteria_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 2 \si{\percent}) alphaproteobacterial sequences on the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater. The proportion of alphaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative (sequences without known relatives within the corresponding group)\label{alpha}}\label{fig:unnamed-chunk-7}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/gammaproteobacteria_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) gammaproteobacterial sequences on the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater. The proportion of gammaproteobacterial sequences in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative (sequences without known relatives within the corresponding group)\label{gamma}}\label{fig:unnamed-chunk-8}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{../results/figures/desulfobacterota_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) sequences within the \textit{Desulfobacterota} on the surfaces of the macrophytes \textit{C. nodosa} (mixed settlement) and \textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater. The proportion of sequences classified as \textit{Desulfobacterota} in the total bacterial and archaeal community is given above the corresponding bar. NR -- No Relative (sequences without known relatives within the corresponding group)\label{desulfo}}\label{fig:unnamed-chunk-9}
\end{figure}

\end{document}