### 2021-07-09 results/numerical/calculator_statistics.tex
Review done.

\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}[t]{>{\centering\arraybackslash}m{6em}ccccc}
\caption{\label{tab:calculator_statistics}Statistic of richness estimator and diversity index parameters for different community types. Parameters were tested by applying the Kruskal-Wallis \textit{H} test followed by a pairwise comparison using the Mann-Whitney \textit{U} test. Bonferroni correction was used to address the problem of multiple comparisons.\label{calculator_statistics}}\\
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{3}{c}{\textbf{Kruskal-Wallis \textit{H} test}} & \multicolumn{2}{c}{\textbf{Mann-Whitney \textit{U} test}} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-6}
\textbf{Parameter} & \textbf{\textit{H}} & \textbf{\textit{df}} & \textbf{\textit{p}} & \textbf{Comparisons Between Community Types} & \textbf{\textit{p}}\\
\midrule
\endfirsthead
\caption[]{Statistic of richness estimator and diversity index parameters for different community types. Parameters were tested by applying the Kruskal-Wallis \textit{H} test followed by a pairwise comparison using the Mann-Whitney \textit{U} test. Bonferroni correction was used to address the problem of multiple comparisons.\label{calculator_statistics} \textit{(continued)}}\\
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{3}{c}{\textbf{Kruskal-Wallis \textit{H} test}} & \multicolumn{2}{c}{\textbf{Mann-Whitney \textit{U} test}} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-6}
\textbf{Parameter} & \textbf{\textit{H}} & \textbf{\textit{df}} & \textbf{\textit{p}} & \textbf{Comparisons Between Community Types} & \textbf{\textit{p}}\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Observed No. of OTUs} & \multirow{-6}{*}{\centering\arraybackslash 44.8} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 1.00\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.01\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Chao1} & \multirow{-6}{*}{\centering\arraybackslash 38.3} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 1.00\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.05\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash ACE} & \multirow{-6}{*}{\centering\arraybackslash 37.1} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 1.00\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & < 0.01\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Exponential Shannon} & \multirow{-6}{*}{\centering\arraybackslash 43.3} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 0.26\\
\cmidrule{1-6}\pagebreak[0]
 &  &  &  & Seawater – \textit{Cymodocea nodosa} (Mixed) & < 0.01\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Mixed) & < 0.0001\\
\nopagebreak
 &  &  &  & Seawater – \textit{Caulerpa cylindracea} (Monospecific) & < 0.0001\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Mixed) & 1.00\\
\nopagebreak
 &  &  &  & \textit{Cymodocea nodosa} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & < 0.01\\
\nopagebreak
\multirow{-6}{6em}{\centering\arraybackslash Inverse Simpson} & \multirow{-6}{*}{\centering\arraybackslash 34.8} & \multirow{-6}{*}{\centering\arraybackslash 3} & \multirow{-6}{*}{\centering\arraybackslash < 0.0001} & \textit{Caulerpa cylindracea} (Mixed) – \textit{Caulerpa cylindracea} (Monospecific) & 0.26\\*
\end{longtable}
\endgroup{}