### 2021-07-09 submission/supplementary.Rmd

---
title: "**Supplementary Material**"
subtitle: "**Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header_supplementary.tex
fontsize: 12pt
geometry: margin=1.0in
csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
bibliography: references.bib
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE, message=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)

library("kableExtra")
```

\vspace{10mm}
Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^

1\. Center for Marine Research, Ruđer Bošković Institute, Croatia

2\. Department of Functional and Evolutionary Ecology, University of Vienna, Austria

3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research (NIOZ), Utrecht University, The Netherlands

^$*$^To whom correspondence should be addressed:

Marino Korlević

G. Paliaga 5, 52210 Rovinj, Croatia

Tel.: +385 52 804 768

Fax: +385 52 804 780

e-mail: marino.korlevic@irb.hr

Running title: Seasonal dynamics of epiphytic communities

[]: # (\linenumbers)
\sisetup{mode=text}
\setlength\parindent{24pt}

## Supplementary figures

```{r out.width="85%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C. nodosa} (mixed settlement) and \\textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{rarefaction}", fig.pos="H"}
include_graphics(normalizePath("results/figures/rarefaction.jpg"))
```

```{r out.width="85%", fig.align="center", fig.cap="Seasonal dynamics of observed number of OTUs, Chao1, ACE, exponential of the Shannon diversity index and Inverse Simpson index of bacterial and archaeal communities from the surfaces of the macrophytes \\textit{C. nodosa} (mixed settlement) and \\textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{calculators}", fig.pos="H"}
include_graphics(normalizePath("results/figures/calculators.jpg"))
```

```{r out.width="85%", fig.align="center", fig.cap="Relative contribution of chloroplast sequences on the surfaces of the macrophytes \\textit{C. nodosa} (mixed settlement) and \\textit{C. cylindracea} (mixed and monospecific settlement) and in the ambient seawater.\\label{chloroplast}", fig.pos="H"}
include_graphics(normalizePath("results/figures/chloroplast_bar_plot.jpg"))
```

## Supplementary table

```{r nseq_notus}
# Loading OTU/sample data
shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")

# Calculating the number of sequences and OTUs in each sample
nseq_notus <- shared %>%
  mutate(nseq=rowSums(select(., starts_with("Otu")))) %>%
  mutate(notus=rowSums(select(., starts_with("Otu"))!=0)) %>%
  select(Group, nseq, notus)

# Loading metadata 
metadata <- read_tsv("data/raw/metadata.csv")

# Joining metadata with the number of sequences and OTUs
invisible(Sys.setlocale(locale="en_GB.utf8"))
metadata_nseq_notus <- inner_join(metadata, nseq_notus, by=c("ID"="Group")) %>%
  select(ID, station, date, season, nseq, notus) %>%
  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
  mutate(date=format(date, "%d %B %Y")) %>%
  mutate(date=str_replace(date, "^0", "")) %>%
  arrange(match(station, c("F", "FCyM", "FCaM", "FCa"))) %>%
  mutate(station=str_replace(station, "^F$", "Seawater")) %>%
  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Mixed)")) %>%
  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
  mutate(ID=str_replace(ID, "^23_1$", "23a")) %>%
  mutate(ID=str_replace(ID, "^23_2$", "23b")) %>%
  mutate(nseq=format(nseq, big.mark=","), notus=format(notus, big.mark=","))

# Generating the table
kable(metadata_nseq_notus, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Sample ID", "Community Type", "Date", "Season", "No. of Sequences", "No. of OTUs"), caption="Sample ID, community type, sampling date and season, no. of sequences and no. of OTUs of each sample. The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\\label{nseq_notus}", align="c") %>%
  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width="6em")
```

```{r calculators}
# Loading rarefied community data created during the generation of the richness and diversity calculators plot
load(file="results/numerical/rarefied.Rdata")

# Copying the sample labels to the rows (input for library vegan)
rarefied <- rarefied %>%
  column_to_rownames("Group")

# Calculating Chao1 and ACE species estimators
estimators <- estimateR(rarefied)
estimators <- as_tibble(t(estimators), .name_repair="unique") %>%
  add_column("Group"=colnames(estimators), .before=TRUE)

# Calculating diversity indices
shannon <- diversity(rarefied, index="shannon")
shannon <- tibble(Group=names(shannon), shannon)
invsimpson <- diversity(rarefied, index="invsimpson")
invsimpson <- tibble(Group=names(invsimpson), invsimpson)

# Transforming the Shannon entropy to the effective number of species
# (http://www.loujost.com/Statistics%20and%20Physics/Diversity%20and%20Similarity/EffectiveNumberOfSpecies.htm)
shannon <- mutate(shannon, shannon=exp(shannon)) %>%
  rename(eshannon=shannon)

# Joining together estimators and indices
estimators_indices <- inner_join(estimators, shannon, by="Group") %>%
  inner_join(., invsimpson, by="Group")

# Loading metadata 
metadata <- read_tsv("data/raw/metadata.csv")

# Joining metadata with estimators and indices
estimators_indices_metadata <- inner_join(metadata, estimators_indices, by=c("ID"="Group")) %>%
  mutate(date=as.Date(date, "%d.%m.%Y"))

# Calculating the mean and standard deviation for each environment
calculators_environemnt <- estimators_indices_metadata %>%
  group_by(station) %>%
  summarise(S.obs_mean=mean(S.obs), S.obs_sd=sd(S.obs),
            S.chao1_mean=mean(S.chao1), S.chao1_sd=sd(S.chao1),
            S.ACE_mean=mean(S.ACE), S.ACE_sd=sd(S.ACE),
            eshannon_mean=mean(eshannon), eshannon_sd=sd(eshannon),
            invsimpson_mean=mean(invsimpson), invsimpson_sd=sd(invsimpson)) %>%
  arrange(match(station, c("F", "FCyM", "FCaM", "FCa"))) %>%
  mutate(station=str_replace(station, "^F$", "Seawater")) %>%
  mutate(station=str_replace(station, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Mixed)")) %>%
  mutate(station=str_replace(station, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
  mutate(station=str_replace(station, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
  mutate_at(vars(2:ncol(.)), ~ format(., big.mark=",", digits=1, nsmall=1, scientific=FALSE)) %>%
  mutate(S.obs=paste(S.obs_mean, " ± ", S.obs_sd),
         S.chao1=paste(S.chao1_mean, " ± ", S.chao1_sd),
         S.ACE=paste(S.ACE_mean, " ± ", S.ACE_sd),
         eshannon=paste(eshannon_mean, " ± ", eshannon_sd),
         invsimpson=paste(invsimpson_mean, " ± ", invsimpson_sd)) %>%
  select(!contains("_"))

# Generating the table
kable(calculators_environemnt, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Community Type", "Observed No. of OTUs", "Chao1", "ACE", "Exponential Shannon", "Inverse Simpson"), caption="Richness estimators and diversity indices of each community type. Calculations were done after normalization to the minimum number of reads per sample.\\label{calculators_community_type}", align="c") %>%
  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width="10em", latex_valign="m") %>%
  column_spec(2, width="10em", latex_valign="m") %>%
  column_spec(5:6, width="6em", latex_valign="m")
```

\newpage
```{r calculator_statistics}
# Performing the Kruskal-Wallis H test
# Observed No. of OTUs
kruskal <- kruskal.test(formula=S.obs ~ station, data=estimators_indices_metadata)
pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$S.obs, estimators_indices_metadata$station,
                                              p.adjust.method="bonferroni", exact=FALSE)
pairwise <- pairwise.wilcox$p.value %>%
  as_tibble() %>%
  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
calculators_environemnt_test <-  tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
                                         "S.obs", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)

# Chao1
kruskal <- kruskal.test(formula=S.chao1 ~ station, data=estimators_indices_metadata)
pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$S.chao1, estimators_indices_metadata$station,
                                              p.adjust.method="bonferroni")
pairwise <- pairwise.wilcox$p.value %>%
  as_tibble() %>%
  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
                     "S.chao1", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)

# ACE
kruskal <- kruskal.test(formula=S.ACE ~ station, data=estimators_indices_metadata)
pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$S.ACE, estimators_indices_metadata$station,
                                              p.adjust.method="bonferroni")
pairwise <- pairwise.wilcox$p.value %>%
  as_tibble() %>%
  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
                     "S.ACE", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)

# Exponential Shannon
kruskal <- kruskal.test(formula=eshannon ~ station, data=estimators_indices_metadata)
pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$eshannon, estimators_indices_metadata$station,
                                              p.adjust.method="bonferroni")
pairwise <- pairwise.wilcox$p.value %>%
  as_tibble() %>%
  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
                     "eshannon", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)

# Inverse Simpson
kruskal <- kruskal.test(formula=invsimpson ~ station, data=estimators_indices_metadata)
pairwise.wilcox <- pairwise.wilcox.test(estimators_indices_metadata$invsimpson, estimators_indices_metadata$station,
                                              p.adjust.method="bonferroni")
pairwise <- pairwise.wilcox$p.value %>%
  as_tibble() %>%
  add_column("V"=rownames(pairwise.wilcox$p.value), .before=TRUE)
test_data <- tribble(~parameter, ~KW_H, ~KW_df, ~KW_p, ~pairwise, ~MW_p,
                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCyM", filter(pairwise, V=="FCyM")$F,
                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCaM", filter(pairwise, V=="FCaM")$F,
                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "F – FCa", filter(pairwise, V=="FCa")$F,
                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCaM", filter(pairwise, V=="FCyM")$FCaM,
                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCyM – FCa", filter(pairwise, V=="FCyM")$FCa,
                     "invsimpson", kruskal$statistic, kruskal$parameter, kruskal$p.value, "FCaM – FCa", filter(pairwise, V=="FCaM")$FCa)
calculators_environemnt_test <- rbind(calculators_environemnt_test, test_data)

# Formatting the statistics table
calculators_environemnt_test <- calculators_environemnt_test %>%
  mutate(pairwise=str_replace(pairwise, "F – FCyM", "Seawater – \\\\textit{Cymodocea nodosa} (Mixed)")) %>%
  mutate(pairwise=str_replace(pairwise, "F – FCaM", "Seawater – \\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
  mutate(pairwise=str_replace(pairwise, "F – FCa", "Seawater – \\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
  mutate(pairwise=str_replace(pairwise, "FCyM – FCaM", "\\\\textit{Cymodocea nodosa} (Mixed) – \\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
  mutate(pairwise=str_replace(pairwise, "FCyM – FCa", "\\\\textit{Cymodocea nodosa} (Mixed) – \\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
  mutate(pairwise=str_replace(pairwise, "FCaM – FCa", "\\\\textit{Caulerpa cylindracea} (Mixed) – \\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
  mutate(KW_p=case_when(KW_p < 0.05 & KW_p >= 0.01 ~ "< 0.05",
                        KW_p < 0.01 & KW_p >= 0.001 ~ "< 0.01",
                        KW_p < 0.001 & KW_p >= 0.0001 ~ "< 0.001",
                        KW_p < 0.0001 ~ "< 0.0001",
                        TRUE ~ as.character(format(round(KW_p, 2), nsmall=2, scientific=FALSE)))) %>%
  mutate(MW_p=case_when(MW_p < 0.05 & MW_p >= 0.01 ~ "< 0.05",
                        MW_p < 0.01 & MW_p >= 0.001 ~ "< 0.01",
                        MW_p < 0.001 & MW_p >= 0.0001 ~ "< 0.001",
                        MW_p < 0.0001 ~ "< 0.0001",
                        TRUE ~ as.character(format(round(MW_p, 2), nsmall=2, scientific=FALSE)))) %>%
  mutate(parameter=str_replace(parameter, "S.obs", "Observed No. of OTUs")) %>%
  mutate(parameter=str_replace(parameter, "S.chao1", "Chao1")) %>%
  mutate(parameter=str_replace(parameter, "S.ACE", "ACE")) %>%
  mutate(parameter=str_replace(parameter, "eshannon", "Exponential Shannon")) %>%
  mutate(parameter=str_replace(parameter, "invsimpson", "Inverse Simpson")) %>%
  mutate(KW_H=format(KW_H, digits=1, nsmall=1))

# Generating the table
calculator_statistics <- kable(calculators_environemnt_test, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Parameter", "\\textit{H}", "\\textit{df}", "\\textit{p}", "Comparisons Between Community Types", "\\textit{p}"), caption="Statistic of richness estimator and diversity index parameters for different community types. Parameters were tested by applying the Kruskal-Wallis \\textit{H} test followed by a pairwise comparison using the Mann-Whitney \\textit{U} test. Bonferroni correction was used to address the problem of multiple comparisons.\\label{calculator_statistics}", align="c") %>%
  add_header_above(c(" ", "Kruskal-Wallis \\\\textit{H} test"=3, "Mann-Whitney \\\\textit{U} test"=2), bold=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width="6em", latex_valign="m") %>%
  collapse_rows(columns=1:2, valign="middle", latex_hline="major")

# Saving the table
cat(calculator_statistics, file="results/numerical/calculator_statistics.tex")

# Merging rows in column H, df and p by parameter
system("sed -i 's/ &  & 3 & < 0.0001 / \\&  \\&  \\&  /' results/numerical/calculator_statistics.tex")
system("sed -i 's/& 3 & < 0.0001 &/\\& \\\\multirow{-6}{*}{\\\\centering\\\\arraybackslash 3} \\& \\\\multirow{-6}{*}{\\\\centering\\\\arraybackslash < 0.0001} \\&/' results/numerical/calculator_statistics.tex")
```
\input{../results/numerical/calculator_statistics.tex}

```{r core_otus_taxonomy}
# Loading rarefied community data created during the generation of the richness and diversity calculators plot
load(file="results/numerical/rarefied.Rdata")

# Joining metadata with OTU/sample data
metadata_rarefied <- read_tsv("data/raw/metadata.csv") %>%
  inner_join(., rarefied, by=c("ID"="Group")) %>%
  mutate(date=as.Date(date, "%d.%m.%Y"))

# Finding the core OTUs of each environment and their taxonomy
# Cymodocea nodosa (mixed)
core <- metadata_rarefied %>%
  filter(station=="FCyM") %>%
  select_if(~ all(. != 0)) %>%
  select(starts_with("Otu"))

total <- metadata_rarefied %>%
  filter(station=="FCyM") %>%
  select(starts_with("Otu"))

core_otus_fcym <- length(core)/length(total) * 100
core_seq_fcym <- sum(core)/sum(total) * 100

core_otu_sum <- as.list(colSums(core)) %>%
  as_tibble() %>%
  gather(key="OTU", value="size_normalized", 1:ncol(.))

otu_taxonomy_fcym <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy") %>%
  inner_join(core_otu_sum, ., by=c("OTU"="OTU")) %>%
  select(-Size)

# Caulerpa cylindracea (mixed)
core <- metadata_rarefied %>%
  filter(station=="FCaM") %>%
  select_if(~ all(. != 0)) %>%
  select(starts_with("Otu"))

total <- metadata_rarefied %>%
  filter(station=="FCaM") %>%
  select(starts_with("Otu"))

core_otus_fcam <- length(core)/length(total) * 100
core_seq_fcam <- sum(core)/sum(total) * 100

core_otu_sum <- as.list(colSums(core)) %>%
  as_tibble() %>%
  gather(key="OTU", value="size_normalized", 1:ncol(.))

otu_taxonomy_fcam <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy") %>%
  inner_join(core_otu_sum, ., by=c("OTU"="OTU")) %>%
  select(-Size)

# Caulerpa cylindracea (monospecific)
core <- metadata_rarefied %>%
  filter(station=="FCa") %>%
  select_if(~ all(. != 0)) %>%
  select(starts_with("Otu"))

total <- metadata_rarefied %>%
  filter(station=="FCa") %>%
  select(starts_with("Otu"))

core_otus_fca <- length(core)/length(total) * 100
core_seq_fca <- sum(core)/sum(total) * 100

core_otu_sum <- as.list(colSums(core)) %>%
  as_tibble() %>%
  gather(key="OTU", value="size_normalized", 1:ncol(.))

otu_taxonomy_fca <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy") %>%
  inner_join(core_otu_sum, ., by=c("OTU"="OTU")) %>%
  select(-Size)
```

\newpage
\blandscape
```{r core_otus_taxonomy_table}
otu_taxonomy_fcym <- otu_taxonomy_fcym %>%
  mutate(environment="\\textit{Cymodocea nodosa} (Mixed)", .before=OTU) %>%
  arrange(desc(size_normalized)) %>%
  slice(1:10)

otu_taxonomy_fcam <- otu_taxonomy_fcam %>%
  mutate(environment="\\textit{Caulerpa cylindracea} (Mixed)", .before=OTU) %>%
  arrange(desc(size_normalized)) %>%
  slice(1:10)

otu_taxonomy_fca <- otu_taxonomy_fca %>%
  mutate(environment="\\textit{Caulerpa cylindracea} (Monospecific)", .before=OTU) %>%
  arrange(desc(size_normalized)) %>%
  slice(1:10)

otu_taxonomy <- rbind(otu_taxonomy_fcym,
                      otu_taxonomy_fcam,
                      otu_taxonomy_fca) %>%
  mutate(Taxonomy=str_replace_all(Taxonomy, ";", "; ")) %>%
  mutate(Taxonomy=str_replace_all(Taxonomy, "\\([^\\)]+\\)", "")) %>%
  mutate(Taxonomy=str_replace_all(Taxonomy, "_", " ")) %>%
  mutate(Taxonomy=str_replace(Taxonomy, "; $", "}")) %>%
  mutate(Taxonomy=str_replace_all(Taxonomy, "^", "\\\\textit{")) %>%
  mutate(Taxonomy=str_replace_all(Taxonomy, "; ", "}; \\\\textit{")) %>%
  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Rhodobacteraceae unclassified\\}", "\\\\textit{Rhodobacteraceae} (NR)")) %>%
  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Pleurocapsa PCC-7319\\}", "\\\\textit{Pleurocapsa}")) %>%
  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{SEEP-SRB4\\}", "SEEP-SRB4")) %>%
  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Desulfocapsaceae unclassified\\}", "\\\\textit{Desulfocapsaceae} (NR)")) %>%
  mutate(Taxonomy=str_replace(Taxonomy, "\\\\textit\\{Desulfobacteraceae unclassified\\}", "\\\\textit{Desulfobacteraceae} (NR)"))

# Generating the table
kable(otu_taxonomy, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Community Type", "OTU Number", "No. of Sequences", "OTU Taxonomy"), caption="Taxonomic classification of OTUs present at every sampling date in the each community type. Only ten OTUs with the highest number of sequences after normalization to the minimum number of reads per sample are shown. NR -- No Relative (sequences without known relatives within the corresponding group)\\label{core_otus_taxonomy_table}", align=c("c", "c", "c", "l")) %>%
  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width="8em", latex_valign="m") %>%
  collapse_rows(columns=1, valign="middle", latex_hline="major")
```
\elandscape